Tissue Enrichment Analysis: TEA

### 2016-03-14 tests/hgf_benchmarking/doc/tea_main_text.tex
added citations to text

\subsection*{Validation of the algorithm and parameter selection}
	In order to select an appropriate dictionary and validate our tool, we found a set of 30 gold standards based on microarray and RNA-seq literature which are believed to be enriched in specific tissues\cite{}. Some of these studies went on to use GFP to identify expression patterns and for this reason we generated a clean Since the expression data is curated from GFP expression at this time and does not include RNA-seq data, these gold standards are statistically independent from the dataset. We wanted to select a dictionary which included enough terms to be specific beyond the largest C. elegans tissues, yet would minimize the number of spurious results and which had a good dynamic range in terms of enrichment fold-change. Selection of a dictionary based only on minimization of spurious results would result in a dictionary with a large number of annotations per tissue, and would therefore include only the major tissues. On the other hand, selecting a dictionary that can detect smaller tissues will bias us towards tissues with lesser annotations. To our knowledge there is no good method for assessing false-positive or false-negative results for annotations.

@@

	%dictionaries that work well
	To select between dictionaries generated between 50, 33 and 25 annotation cut-offs, and also to ensure the terms that are selected as enriched by our algorithm are reasonable, we looked in detail at the enrichment analysis results. 
	%Work well
	Most results were highly comparable and in line with what was expected. For some sets, all dictionaries seemed to perform well. For example, in our `all neuron enriched set'~\ref{WBPaper00037950} the result was an amalgamation of neuron related terms including mechanosensory neurons, thermosensitive neurons, interneurons, ganglions and male rays regardless of the dictionary used. On the other hand, when we looked at a gene set enriched for germline precursor expression in the embryo~\ref{WBPaper00037950}, the dictionary with the 50 cutoff was only able to identify `oocyte WBbt:006797'; whereas the two smaller dictionaries were able to single out cells germline precursor cells -- at the 33-cutoff, our tool identified `Z2' and `Z3' as being five-fold enriched; whereas at the 25 gene-cutoff the terms `Psub4',`Psub3' and `Psub2' were identified in addition to `Z2' and `Z3'.
	%Differences between dictionaries
	We queried an embryonic stage intestine precursor associate geneset~\ref{WBPaper00037950}. Notably, this gene set yielded no enrichment when using the 25 cutoff dictionary, nor when using the 50 cutoff dictionary. However, the 33 cutoff dictionary suggested, probably correctly, that the E lineage was heavily enriched in this set. 
	%Bad stuff.
	Not all queries worked equally well. For example, a number of intestinal enriched genes sets~\ref{WBPaper00026980, WBPaper00037950} were not enriched in intestine in any dictionary, but they were enriched for pharynx- and hypodermis-related terms. We were somewhat surprised that intestinal gene sets performed poorly, since the intestine is a relatively well-annotated tissue.
	%Dictionary intra-agreement
	We also assessed the internal agreement of our tool by using independent gene-sets that we expected to be enriched in the same tissues. We had two independent pan-neuronal sets~\ref{}; two independent PVD enriched sets~\ref{}; two independent GABAergic gene sets~\ref{}; two independent pharyngeal gene sets; and two independent intestinal gene sets~\ref{}. Overall, the tool seems to have good internal agreement. On most sets, the same terms were enriched, although order was somewhat variable. However, most high-scoring terms were preserved between gene sets. The intestinal gene-sets and pharyngeal gene sets comparisons were exceptions, since at least one gene set was missing each for intestine and pharynx in every dictionary, so we didn't consider them as informative for assessing internal agreement.

	 All comparisons can be found online at: www.XXX.com. Overall, the dictionary generated by a 33 gene annotation cutoff with 0.95 redundancy threshold using the `any' criterion. seemed to perform well, with a good balance between specificity, verbosity and accuracy, so we selected this parameter set to generate our static dictionary.

@@

>>>>

\subsection*{Validation of the algorithm and parameter selection}
	In order to select an appropriate dictionary and validate our tool, we found a set of 30 gold standards based on microarray and RNA-seq literature which are believed to be enriched in specific tissues\cite{Gaudet2004, Spencer2011, Cinar2005, Watson2008a, Pauli2006, Portman2004, Fox2007, Smith2010}. Some of these studies went on to use GFP to identify expression patterns and for this reason we generated a clean Since the expression data is curated from GFP expression at this time and does not include RNA-seq data, these gold standards are statistically independent from the dataset. We wanted to select a dictionary which included enough terms to be specific beyond the largest C. elegans tissues, yet would minimize the number of spurious results and which had a good dynamic range in terms of enrichment fold-change. Selection of a dictionary based only on minimization of spurious results would result in a dictionary with a large number of annotations per tissue, and would therefore include only the major tissues. On the other hand, selecting a dictionary that can detect smaller tissues will bias us towards tissues with lesser annotations. To our knowledge there is no good method for assessing false-positive or false-negative results for annotations.

@@

	%dictionaries that work well
	To select between dictionaries generated between 50, 33 and 25 annotation cut-offs, and also to ensure the terms that are selected as enriched by our algorithm are reasonable, we looked in detail at the enrichment analysis results. 
	%Work well
	Most results were highly comparable and in line with what was expected. For some sets, all dictionaries seemed to perform well. For example, in our `all neuron enriched sets'~\cite{Spencer2011, Watson2008a} the results were an amalgamation of neuron related terms including mechanosensory neurons, thermosensitive neurons, interneurons, ganglions and male rays regardless of the dictionary used. On the other hand, when we looked at a gene set enriched for germline precursor expression in the embryo~\cite{Spencer2011}, the dictionary with the 50 cutoff was only able to identify `oocyte WBbt:006797'; whereas the two smaller dictionaries were able to single out cells germline precursor cells -- at the 33-cutoff, our tool identified `Z2' and `Z3' as being five-fold enriched; whereas at the 25 gene-cutoff the terms `Psub4',`Psub3' and `Psub2' were identified in addition to `Z2' and `Z3'.
	%Differences between dictionaries
	We queried an embryonic stage intestine precursor geneset~\cite{Spencer2011}. Notably, this gene set yielded no enrichment when using the 25 cutoff dictionary, nor when using the 50 cutoff dictionary. However, the 33 cutoff dictionary suggested, probably correctly, that the E lineage was heavily enriched in this set. 
	%Bad stuff.
	Not all queries worked equally well. For example, a number of intestinal enriched genes sets~\cite{Spencer2011, Pauli2006} were not enriched in intestine in any dictionary, but they were enriched for pharynx- and hypodermis-related terms. We were somewhat surprised that intestinal gene sets performed poorly, since the intestine is a relatively well-annotated tissue.
	%Dictionary intra-agreement
	We also assessed the internal agreement of our tool by using independent gene-sets that we expected to be enriched in the same tissues. We had two independent pan-neuronal sets~\cite{Spencer2011, Watson2008a}; two independent PVD enriched sets~\cite{Spencer2011, Smith2010}; two independent GABAergic gene sets~\cite{Spencer2011, Cinar2005}; two independent pharyngeal gene sets~\cite{Spencer2011, Gaudet2004}; and two independent intestinal gene sets~\cite{Spencer2011, Pauli2006}. Overall, the tool seems to have good internal agreement. On most sets, the same terms were enriched, although order was somewhat variable. However, most high-scoring terms were preserved between gene sets. The intestinal gene-sets and pharyngeal gene sets comparisons were exceptions, since at least one gene set was missing each for intestine and pharynx in every dictionary, so we didn't consider them as informative for assessing internal agreement. 
	 All comparisons can be found online in our Github repository (see Availability of data and materials). Overall, the dictionary generated by a 33 gene annotation cutoff with 0.95 redundancy threshold using the `any' criterion. seemed to perform well, with a good balance between specificity, verbosity and accuracy, so we selected this parameter set to generate our static dictionary.

@@

\section*{Availability of data and materials}
Our web implementation is available at https://www.wormbase.org/tea. Our software can also be downloaded using Python's pip installer via the command

\texttt{pip install tissue\_enrichment\_tool}

Alternatively, our software is available for download at:

http://dangeles.github.io/TissueEnrichmentAnalysis.

All benchmark gene sets, benchmarking code and figures can also be found at the same address, under the `tests' folder.
